Topological order in the quantum dimer-pentamer model

### 2016-04-13 qdpm_paper.tex
first commit in a while. Started refining the text and equations.

    Now we will introduce the toric code on the square lattice in some detail to sever as a point of reference and
    comparison for the following discussion of the $Z_3$ case. The toric code Hamiltonian is given by
    %
    \begin{equation}
        H_t = -J_e\sum_v A_v - J_m\sum_p B_p
        \label{eqn:toric_code_ham}
    \end{equation}
    %
    where the operator $A_v$ acts on the vertex $v$ and the sum over $v$ is the sum over all vertices. 
    The operator $B_p$ acts on a plaquette $p$ and the sum over $p$ is the sum over all
    plaquettes. In the $Z_2$ toric code, Ising degrees of freedom live on the links of the
    lattice. Explicitly the operators $A_v$ and $B_p$ can be expressed as
    %
    \begin{equation}
        A_v = \prod_{i\in v} \sigma^x_i
        ,
    \end{equation}
    %
    and
    \begin{equation}
        B_p = \prod_{i\in p} \sigma^z_i
        .
    \end{equation}
    %
    The products are over the links adjacent to vertex $v$ and on the edges of plaquette $p$ for
    $A_v$ and $B_p$ respectively. Since all $A_v$ operators commute with each other, all
    $B_p$ operators commute with each other and $A_v$ commutes with $B_p$ even if they share
    links (commute because they
    can only share an even number of links) then the Hamiltonian can be solved term by term.

    Without a rigorous solution it is possible to guess at the form of the Toric code
    ground state wave function. The first term in \ref{eqn:toric_code_ham} is maximized ($H_t$
    minimized) when an even number of links touching a vertex are spin up or down in the
    $\sigma^z$ basis. By coloring the links of all down spins the groundstate will
    contain closed loops. The second term in \ref{eqn:toric_code_ham} is maximized for a
    superposition of loops since $B_p$ flips the spins around a plaquette $p$. The ground state
    wave function is then a superposition of all possible closed loop configurations. 
    %Since $B_p$ is a local operator
    %though the alowed loop configurations are determined by how the sytem in initilized.

    NEED HERE Some stuff about plaquette flux ?? -> $B_p | \Psi \rangle = | \Psi \rangle$ -> only true if
    grnd state as no vortices

    NEED HERE String operators. The electric and magnetic string operators are respectively
    \begin{equation}
        S^e_{\Gamma} = \prod_{i\in\Gamma_e} \sigma^x_i
        ,
    \end{equation}
    \begin{equation}
        S^m_{\Gamma} = \prod_{i\in\Gamma_m} \sigma^z_i
        ,
    \end{equation}
    where the path is denoted by $\Gamma$. The path of the electric string operator is along the
    links of the lattice and if this path is not closed then the operator
    creates a pair of excitations. These excitations are defects in the system 
    that can be thought of as electric charges that live at the
    end of the unclosed string. Each of these electric defects has an energy costs of $2J_e$. 
    The path of the magnetic string operator runs perpendicular to the links
    and starts and ends on the dual lattice. An unclosed magnetic path results in two magnetic
    excitations on the at the ends of the magnetic string each costing $2J_m$ each.

    Using a nontrivial path (genus of system $>0$) one can move between topological sectors using $S_e$ and determine
    the topological sector the system is in using $S_m$. On a torus, acting with $S_e$ along a
    path that encloses either the major or minor axis of the torus changes the topological sector.
    In general there are $4^g$ topological sectors where $g$ is the
    genus of the system. For a torus then there are four topological sectors. If These can be
    measuerd using the $S_m$ operator on a ...

    \subsection{$Z_3$ Toric Code}
        We can represent the $Z_3$ algebra through the following expressions
        %
        %Z_N
        %\begin{equation}
        %    E |n\rangle = n |n\rangle, ~ n=0,1,2,...,N-1
        %    \\
        %    e^{iA} |n\rangle = |n+1 \rangle, ~ n=0,1,2,...,N-2
        %    \\
        %    e^{iA} |N-1\rangle = |0\rangle
        %    %\label{eqn:}
        %\end{equation}
        %Z_3
        %
        \begin{equation}
            \begin{split}
            & E_l |n_l\rangle = n_l |n_l\rangle, ~ n_l=0,1,2
            \\
            & e^{iA_l} |n_l \rangle = |n_l+1 \rangle, ~ n_l =0,1
            \\ 
            & e^{iA_l} |2 \rangle = |0\rangle
            \end{split}
            ,
            %\label{eqn:}
        \end{equation}
        %
        where the operators $E_l$ and $A_l$ act on a given link $l$ and are analogous to the electric field and vector potential
        respectively. 
        By assigning orientations to the links between two vertices we can define 3 unique
        link ``values'' pictorially. 
        %
        \begin{itemize}
            \item $\vvemptylink$ 
            \item $\rightarrowlink$ 
            \item $\leftarrowlink$ 
        \end{itemize}
        % 
        In this way we can define $e^{iA_l}$ as the operator $Q$ which acts on the links in the
        following way
        \begin{equation}
            \begin{split}
                & Q^\dagger_{vv'}  | \vvemptylink \rangle = | \rightarrowlink \rangle
                ,
                \\
                & Q^\dagger_{vv'}  | \rightarrowlink \rangle = | \leftarrowlink \rangle
                ,
                \\
                & Q^\dagger_{vv'}  | \leftarrowlink \rangle = | \emptylink \rangle
                ,
            \end{split}
            %\label{eqn:}
        \end{equation}
        where the link is specified by the two vertices it connects and the operator
        $Q^\dagger_{vv'}$ acts in a direction from vertex $v$ to vertex $v'$. By changing the
        direction in which $Q$ operates we can express it in terms of its hermitian conjugate. 
        %
        \begin{equation}
            (Q^\dagger_{vv'})^\dagger = Q_{vv'}^\dagger
            .
            %\label{eqn:}
        \end{equation}

        The operator $E$ The operator $E_{vv'}$ can be defined in a similar way as
        %
        \begin{equation}
            \begin{split}
                & E^\dagger_{vv'}  | \vvemptylink \rangle = 0
                ,
                \\
                & E^\dagger_{vv'}  | \rightarrowlink \rangle = | \rightarrowlink \rangle
                ,
                \\
                & E^\dagger_{vv'}  | \leftarrowlink \rangle = 2| \leftarrowlink \rangle
                ,
            \end{split}
            %\label{eqn:}
        \end{equation}
        and with it we can define.
        %
        \begin{equation}
            P_{vv'}^\dagger \equiv e^{i2\pi E_{vv'}/3}
            .
            %\label{eqn:}
        \end{equation}
        %
        Now how $P_{vv'}^\dagger$ acts on the links: 
        %
        \begin{equation}
            \begin{split}
                & P^\dagger_{vv'}  | \vvemptylink \rangle = | \vvemptylink \rangle
                ,
                \\
                & P^\dagger_{vv'}  | \rightarrowlink \rangle = e^{i2\pi/3} | \rightarrowlink \rangle
                ,
                \\
                & P^\dagger_{vv'}  | \leftarrowlink \rangle = e^{-i2\pi/3} | \leftarrowlink \rangle
                .
            \end{split}
            %\label{eqn:}
        \end{equation}
        %

        The $Z_3$ toric code Hamiltonian can now be written in terms of the $P$ and $Q$ operators. The
        usual string net Hamiltonian is written as 
        \begin{equation}
            H = - J_e \sum_v A_v - J_m \sum_p B_p
            %\label{eqn:}
        \end{equation}
        where the sum over $v$ and the sum over $p$ are the sums over all vertices and plaquettes on
        the lattice respectively and 
        \begin{equation}
            A_v = \prod_{i\in v} P^\dagger_{vv_i}~,~ B_p \prod_{i\in p} Q^\dagger_{v_iv_{i+1}}
            .
            %\label{eqn:}
        \end{equation}
        The specific vertex and plaquette labeling scheme is shown in Fig.~\ref{fig:vertex_link_labels}.
        %
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.8\linewidth]{vertex_link_gauge_def}
            \caption{Notation of vertex and plaquette labels}
            \label{fig:vertex_link_labels}
        \end{figure}
        %
        For the ground state of this Hamiltonian there is a local symmetry corresponding to the
        minimization of the first term. The model has an exactly local gauge symmetry that can be
        expressed as
        \begin{equation}
            G_v | \psi \rangle= | \psi \rangle
            %\label{eqn:}
        \end{equation}
        where $G_v$ = $A_v$.

        \subsubsection{Hamiltonian Gauge symmetry }

        \subsubsection{String operators}
            With the above notation we can write down the string operators. The electric string operator along
            the path $\Gamma_{v_b v_e}$ between the two vertices $v_0$ (beginning) and $v_N$ (end) is written
            %
            \begin{equation}
                S^e_{\Gamma} = \prod_{l=0}^{N-2} Q_{v_lv_{l+1}}^\dagger
                %\label{eqn:}
                ,
            \end{equation}
            %
            where $v_l \in \Gamma$ and $N$ is the number of vertices in $\Gamma$. We show an example of an
            electric string operator operation on an example configuration in Fig.~\ref{fig:example_elec_string}.
            %
            %
            \begin{figure}[htpb]
                \centering
                \includegraphics[width=0.8\linewidth]{example_elec_string.pdf}
                \caption{The grey line is the path $\Gamma$ on which the string operator acts. The operator acts
            in the direction from the vertex labeled in red to the one labeled in blue.}
                \label{fig:example_elec_string}
            \end{figure}
            %
            %
            The magnetic string operator can be written as
            \begin{equation}
                S^m_{\Gamma} = \prod_{l=0}^{N-1} Q_{v_{2l}v_{2l+1}}^\dagger
                %\label{eqn:}
            \end{equation}
            %
            where $v_l$ are the vertices adjacent to the links in $\Gamma$ and $N$ is the number of
            vertices
            on the \textit{left} side of the path when looking along the path from red to blue. 
            An example of this operator acting
            on a configuration is shown in Fig.~\ref{fig:example_mag_string} along with the vertex labeling
            scheme.
            %
            %
            \begin{figure}[htpb]
                \centering
                \includegraphics[width=0.8\linewidth]{example_mag_string.pdf}
                \caption{The grey line is the path $\Gamma$ on which the string operator acts. The magnetic
                    string operator acts in the 
                    direction from the plaquette labeled in red to the one labeled in blue.}
                \label{fig:example_mag_string}
            \end{figure}
            %
            The magnetic string operator acting on path $\Gamma$ in Fig.~\ref{fig:example_mag_string} is
            explicitly
            %
            \begin{multline}
                S^m_{\Gamma} (\mathrm{~Fig.~\ref{fig:example_mag_string}~config})
                = P_{v_0v_1}^\dagger P_{v_2v_3}^\dagger P_{v_4v_5}^\dagger
                \\
                = \exp{[0 - \frac{i2\pi}{3} + \frac{i2\pi}{3} ]} 
                = 1 (\mathrm{~Fig.~\ref{fig:example_mag_string}~config})
                .
            \end{multline}
            %

            An interesting property of the electric string string operators in the $Z_3$ toric code is
            that they produce two defects, one having positive charge and the other having negative
            charge. This is most easily see by looking at Fig.~\ref{fig:example_elec_string}. The
            configuration in Fig.~\ref{fig:example_elec_string}(a) obeys the local gauge symmetry and
            when the electric string operator acts on path $\Gamma$ it results in two defects at the
            end points of $\Gamma$. At the first $v_0$ (red) the local constraint is broken and
            $G_{v_0} = e^{i2\pi /3}$ as shown in Fig.~\ref{fig:example_elec_string}(b), whereas
            $G_{v_3} = e^{-i2\pi /3}$ which can be thought of as negative and positive gauge charges
            at $v_0$ and $v_3$ respectively.

            The magnetic excitations, or $Z_3$ visons live at the endpoints of a path created with the magnetic
            string operator. These visons can take three different values which can be changed by
            operating with the electric operator on a path that encircles one endpoint of the
            magnetic string path.

        \subsubsection{properties of $Z_3$ string operators}

@@

            The ground state of the $Z_3$ toric code on the torus is nine-fold degenerate because of
            the nine different topological sectors in this topology. One can move between the
            topological sectors with a nontrivial closed electric string wraps around either
            the major axis or the minor axis of the torus. We will call the electric winding
            operator that acts around the major axis $W^e_y$ and around the minor axis $W^e_x$.
            A given topological sector can be distinguish 
            by acting around the minor and major axes with the magnetic string operator $W^m_y$ and
            $W^m_x$. All of the aforementioned winding operators commute with the Hamiltonian and
            therefore share the same eigenvalues. The operator $W^m_{x}$ ($W^m_{y}$) can 
            have three different values depending on how many electric strings enclose the minor
            (major) axis. Since there are three different
            possible values for each winding number around each axis there are nine distinct
            topological sectors.

@@

    \subsection{QDM on Square lattice}
        Here we will review the relevant
        properties and attributes of the QDM on the square lattice for which the Hamiltonian is.
        \begin{equation}
            \label{}
            H_{\mathrm{QDM}} = -t\sum_{\bplqt} 
                \left(
                    \left|
                        \vrtplqt
                    \right\rangle
                    \left\langle
                        \hrzplqt
                    \right|
                    +
                    \mathrm{h.c.}
                \right)
            - v\sum_{\bplqt}
                \left(
                    \left|
                        \vrtplqt
                    \right\rangle
                    \left\langle
                        \vrtplqt
                    \right|
                    +
                    \left|
                        \hrzplqt{}
                    \right\rangle
                    \left\langle
                        \hrzplqt
                    \right|
                \right)
                .
        \end{equation}
        This model can be defined in terms of a single parameter $t/v$ for which the wave function is
        known exactly at the Rokhsar and Kivelson (RK) point, the point where $t/v =1$. At the RK
        point the wave function is the equal weighted superposition over all dimer configurations.
        HERE

@@

    In this section we discuss the dimer dimer correlation function
    \begin{equation}
        C_{\mathrm{dimer}} = \langle d_0 d_r \rangle - \langle d_0 \rangle   \langle d_r \rangle   
        ,
    \end{equation}
    where $d_i$ is the ising degrees of freedom on link i.
    In order for the QDPM to exhibit topological order there we must find that it is in a gaped
    dimer liquid phase. It is likely that the QDPM in a liquid phase at the RK point because there 
    we find no evedence of broken symmetries in the system sizes up to $32x32$. As an example of the
    absence of broken symmetries we show
    the dimer correlation function measuered between a single dimer and all other dimers on the
    lattice in Fig.~\ref{fig:fig:dimer_gry_scale_qdpm} for a $16x16$ lattice.
    \subsection{Liquid}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{dimer_gry_scale_qdpm.pdf}
        \caption{Gray scale (actual data) of the dimer corelation between the dimer in the upper
        left corner (colored red) and all other dimers on a $16\times16$ lattice. The 
        dimers in the upperleft corner are colored red because they are off the scale shown at the
        botom of the figure.}
        \label{fig:dimer_gry_scale_qdpm}
    \end{figure}

    We also find exponetialy decaying dimer correlations 
    In Fig.~\ref{fig:fig:spacial_dmr_cor} we show the dimer correlations between horizontal dimers
    in a vertical column. We find the correlation length to be $HERE$.
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{spacial_dmr_cor.pdf}
        \caption{(Note: This figure was made using equal probabilities of plaquette flips and
        pentamer moves) Spacial correlation function of parallel dimers for the two different
        sublattices distinguished by their color (red, green).}
        \label{fig:spacial_dmr_cor}
    \end{figure}

    \subsection{Comparison to other models}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating the Gap}
    The exponetial decay of the spacial correlations shown in the previous sections 
    suggest a gap in the spectrum though it is important to determine if the lowest lying
    excitations are in the vison-like
    or the dimer-like.  The energy gap of the dimer-like and vison-like excitations can be estimated 
    by investigating the dimer and
    vison imaginary time correlations respectivly as outlined by Henley \cite{Henley2004a}. 
    One changlenge in this type of gap estimate is in resloving the correlation functions far enough out
    in imaginary time. In appendix \ref{appendix_gap_err} we discuss our methods used to determine
    that this particular source of systematic error is sufficinetly comparied to the gap
    estimates for different types of excitatoins.

>>>>

    Now we will introduce the toric code on the square lattice which will sever as a reference and
    point of comparison to the discussion of the $Z_3$ case. The toric code Hamiltonian is given by
    %
    \begin{equation}
        H_t = -J_e\sum_v A_v - J_m\sum_p B_p
        \label{eqn:toric_code_ham}
    \end{equation}
    %
    where the operator $A_v$ ($B_p$) acts on the vertex $v$ (plaquette $p$), and the sum over $v$ ($p$) 
    is the sum over all vertices (plaquettes). 
    In the toric code, Ising degrees of freedom live on the links of the
    lattice and the operators $A_v$ and $B_p$ can be expressed as
    %
    \begin{equation}
        A_v = \prod_{i\in v} \sigma^x_i
        ,
    \end{equation}
    %
    and
    \begin{equation}
        B_p = \prod_{i\in p} \sigma^z_i
        .
    \end{equation}
    %
    The products are over the links touching vertex $v$ and the edges of plaquette $p$ for
    $A_v$ and $B_p$ respectively. It is not difficult to show that 
    $[A_v,A_{v'}]=0$ for all $v$ and $v'$, 
    $[B_p,B_{p'}]=0$ for all $p$ and $p'$, and $[A_v,B_p]=0$. The comutation relation between $A$
    and $B$ holds even if they share links because they
    can only share an even number of links. Given the commutation relations,
    the toric Hamiltonian can be solved term by term.

    Even without a rigorous solution it is possible to guess at the form of the Toric code
    ground state wave function. The first term in \ref{eqn:toric_code_ham} is maximized 
    when an even number of links touching each vertex are spin up or down in the
    $\sigma^z$ basis. By coloring the links of all down spins the groundstate will
    contain closed loops. The second term in \ref{eqn:toric_code_ham} is maximized for a
    superposition of loops since $B_p$ flips the spins around a plaquette $p$. The ground state
    wave function is then a superposition of all possible closed loop configurations. 
    %Since $B_p$ is a local operator
    %though the alowed loop configurations are determined by how the sytem in initilized.

    The electric and magnetic string operators are
    \begin{equation}
        S^e_{\Gamma} = \prod_{i\in\Gamma_e} \sigma^x_i
        ,
    \end{equation}
    \begin{equation}
        S^m_{\Gamma} = \prod_{i\in\Gamma_m} \sigma^z_i
        ,
    \end{equation}
    where the path is denoted by $\Gamma$. The electric string operator acts along the
    links whereas the magnetic string operator acts along a path that perpendicularly
    intersects the links. 
    Electric and magnetic excitations are created when the string operator acts on an unclosed path. 
    The electric excitations create monomer defects in the lattice at the end of the string, 
    each of defects costing energy of $2J_e$. 
    The magnetic excitations, which are 
    are variational excited states  orthogonal
    to the groundstate, create excitations at the ends in the ceter of two plaquettes.
    Each end of the magnetic
    string, a vison, costs energy of $2J_m$.

    operating on a nontrivial path (genus of system $>0$) with $S_e$ changes the topological sector.
    The current topological sector can be determined
    by acting around the congugate path with $S_m$, e.g. on a torus, acting with $S_e$ along a
    path that encloses the major or minor axis changes the topological sector and acting on a path
    that encloses the minor axis with $S_m$ measures the topological sector.
    As defined the magnetic string can only produce two different values called winding number.
    Therefore, there are $4^g$ topological sectors on the torus where $g$ is the
    genus of the system.

    \subsection{$Z_3$ Toric Code on the Square Lattice}
        In the $Z_3$ toric three degrees of freedom live on each link.
        We can represent the $Z_3$ algebra through the following expressions
        %
        %Z_N
        %\begin{equation}
        %    E |n\rangle = n |n\rangle, ~ n=0,1,2,...,N-1
        %    \\
        %    e^{iA} |n\rangle = |n+1 \rangle, ~ n=0,1,2,...,N-2
        %    \\
        %    e^{iA} |N-1\rangle = |0\rangle
        %    %\label{eqn:}
        %\end{equation}
        %Z_3
        %
        \begin{equation}
            \begin{split}
            & E_l |n_l\rangle = n_l |n_l\rangle, ~ n_l=0,1,2
            \\
            & e^{iA_l} |n_l \rangle = |n_l+1 \rangle, ~ n_l =0,1
            \\ 
            & e^{iA_l} |2 \rangle = |0\rangle
            \end{split}
            ,
            %\label{eqn:}
        \end{equation}
        %
        where the operators $E_l$ and $A_l$ act on a given link $l$ and are analogous to the 
        electric field and vector potential respectively. 
        By assigning orientations to the links between two vertices we can define the three degrees of
        freedom pictorially. 
        %
        \begin{itemize}
            \item $\vvemptylink$ 
            \item $\rightarrowlink$ 
            \item $\leftarrowlink$ 
        \end{itemize}
        % 
        To simplify notation, we define $e^{iA_l}$ as the operator $Q$ which acts on the links in the
        following way
        \begin{equation}
            \begin{split}
                & Q^\dagger_{vv'}  | \vvemptylink \rangle = | \rightarrowlink \rangle
                ,
                \\
                & Q^\dagger_{vv'}  | \rightarrowlink \rangle = | \leftarrowlink \rangle
                ,
                \\
                & Q^\dagger_{vv'}  | \leftarrowlink \rangle = | \emptylink \rangle
                .
            \end{split}
            %\label{eqn:}
        \end{equation}
        The link is specified by the two vertices it connects. The operator
        $Q^\dagger_{vv'}$ acts in the direction from the first subscript to the second subscript, in
        this case denoted as $v$ and $v'$. 
        Changing the direction in which $Q$ operates is equivalent to taking the hermitian conjugate
        of the operator withougt changing the direction.  
        %
        \begin{equation}
            (Q^\dagger_{vv'})^\dagger = Q_{vv'}^\dagger
            .
            %\label{eqn:}
        \end{equation}

        We now define an operator $E_{vv'}$ in a similar way, 
        %
        \begin{equation}
            \begin{split}
                & E^\dagger_{vv'}  | \vvemptylink \rangle = 0
                ,
                \\
                & E^\dagger_{vv'}  | \rightarrowlink \rangle = | \rightarrowlink \rangle
                ,
                \\
                & E^\dagger_{vv'}  | \leftarrowlink \rangle = 2| \leftarrowlink \rangle
                ,
            \end{split}
            %\label{eqn:}
        \end{equation}
        and with it define the last important operator that acts on the links. 
        %
        \begin{equation}
            P_{vv'}^\dagger \equiv e^{i2\pi E_{vv'}/3}
            .
            %\label{eqn:}
        \end{equation}
        %
        $P_{vv'}^\dagger$ acts on the links in the following way: 
        %
        \begin{equation}
            \begin{split}
                & P^\dagger_{vv'}  | \vvemptylink \rangle = | \vvemptylink \rangle
                ,
                \\
                & P^\dagger_{vv'}  | \rightarrowlink \rangle = e^{i2\pi/3} | \rightarrowlink \rangle
                ,
                \\
                & P^\dagger_{vv'}  | \leftarrowlink \rangle = e^{-i2\pi/3} | \leftarrowlink \rangle
                .
            \end{split}
            %\label{eqn:}
        \end{equation}
        %

        With the above definitions the $Z_3$ toric code Hamiltonian can be written in terms of the $P$ and $Q$ operators.
        In general a string net Hamiltonian might be written as 
        \begin{equation}
            H = - J_e \sum_v A_v - J_m \sum_p B_p
            %\label{eqn:}
        \end{equation}
        where 
        \begin{equation}
            A_v = \prod_{i\in v} P^\dagger_{vv_i}~,~ B_p = \prod_{i\in p} Q^\dagger_{v_iv_{i+1}}
            .
            %\label{eqn:}
        \end{equation}
        The specific vertex and plaquette labeling scheme is shown in Fig.~\ref{fig:vertex_link_labels}.
        %
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.8\linewidth]{vertex_link_gauge_def}
            \caption{Notation of vertex and plaquette labels}
            \label{fig:vertex_link_labels}
        \end{figure}
        %
        For the ground state of this Hamiltonian there is aa local symmetry corresponding to the
        minimization of the first term. This local symmetry is a gauge symmetry that can be
        expressed in terms of the gauge operator.
        \begin{equation}
            G_v | \psi \rangle= | \psi \rangle
            %\label{eqn:}
        \end{equation}
        where $G_v$ = $A_v$.

        \subsubsection{$Z_3$ String operators}
            With the above notation we can now write down the $Z_3$ string operators. The electric string operator along
            the path $\Gamma$ is written
            %
            \begin{equation}
                S^e_{\Gamma} = \prod_{l=0}^{N-2} Q_{v_lv_{l+1}}^\dagger
                %\label{eqn:}
                ,
            \end{equation}
            %
            where $v_l \in \Gamma$ and $N$ is the number of vertices in $\Gamma$. We show an example of an
            electric string operator operatiing on an example configuration in Fig.~\ref{fig:example_elec_string}.
            %
            %
            \begin{figure}[htpb]
                \centering
                \includegraphics[width=0.8\linewidth]{example_elec_string.pdf}
                \caption{The grey line is the path $\Gamma$ on which the string operator acts. The operator acts
            in the direction from the vertex labeled in red to the one labeled in blue.}
                \label{fig:example_elec_string}
            \end{figure}
            %
            %
            The magnetic string operator can be written as
            \begin{equation}
                S^m_{\Gamma} = \prod_{l=0}^{N-1} P_{v_{2l}v_{2l+1}}^\dagger
                %\label{eqn:}
            \end{equation}
            %
            where $v_l$ are the vertices adjacent to the links in $\Gamma$ and $N$ is the number of
            vertices
            on the \textit{left} side of the path when looking along the path from red to blue. 
            An example of this operator acting
            on a configuration is shown in Fig.~\ref{fig:example_mag_string} along with the vertex labeling
            scheme.
            %
            %
            \begin{figure}[htpb]
                \centering
                \includegraphics[width=0.8\linewidth]{example_mag_string.pdf}
                \caption{The grey line is the path $\Gamma$ on which the string operator acts. The magnetic
                    string operator acts in the 
                    direction from the plaquette labeled in red to the one labeled in blue.}
                \label{fig:example_mag_string}
            \end{figure}
            %
            The magnetic string operator acting on path $\Gamma$ in Fig.~\ref{fig:example_mag_string} is
            %
            \begin{multline}
                S^m_{\Gamma} (\mathrm{~Fig.~\ref{fig:example_mag_string}~config})
                = P_{v_0v_1}^\dagger P_{v_2v_3}^\dagger P_{v_4v_5}^\dagger
                \\
                = \exp{[0 - \frac{i2\pi}{3} + \frac{i2\pi}{3} ]} 
                = 1 (\mathrm{~Fig.~\ref{fig:example_mag_string}~config})
                .
            \end{multline}
            %

            An interesting property of the electric string operator in the $Z_3$ toric code is
            that an unclosed string produces two different defects. 
            An example of an open string and the defects is shown in Fig.~\ref{fig:example_elec_string}. The
            configuration in Fig.~\ref{fig:example_elec_string}(a) obeys the local gauge symmetry
            aside from the endpoints of the path that it acts on ($\Gamma$). The two defects at the
            end points produce two different values when the vertecies are operated on by $G_v$, being
            $G_{\mathrm{red}} = e^{i2\pi /3}$, and $G_{\mathrm{blue}} = e^{-i2\pi /3}$.

            The magnetic excitations, or $Z_3$ visons live at the endpoints of the magnetic
            strings.  These visons can change the phase of a configuration by different values.

%        \subsubsection{properties of $Z_3$ string operators}

@@

            The ground state of the $Z_3$ toric code has nine-fold topological degeneracy because of
            the $3\times3$ different topological sectors. As the case of the toric code the $Z_3$
            electric string changes the topological sector when acting on on a nontrivial closed
            path. We call the electric winding
            operator that acts around the major axis $W^e_y$ and around the minor axis $W^e_x$.
            The topological sector can be determined 
            by acting around the axes with the magnetic string operator, called $W^m_y$ and
            $W^m_x$. All of the winding commute with the Hamiltonian and
            therefore share the same eigenvalues. The operator $W^m_{x}$ ($W^m_{y}$) can 
            have three different values depending on how many electric strings enclose the minor
            (major) axis. The three different
            possible values for each winding number around each axis gives the model nine distinct
            topological sectors.

@@

    \subsection{QDM on Square lattice}
        Here we will review the relevant
        properties and attributes of the QDM on the square lattice for which the Hamiltonian is.
        \begin{equation}
            \label{}
            H_{\mathrm{QDM}} = -t\sum_{\bplqt} 
                \left(
                    \left|
                        \vrtplqt
                    \right\rangle
                    \left\langle
                        \hrzplqt
                    \right|
                    +
                    \mathrm{h.c.}
                \right)
            - v\sum_{\bplqt}
                \left(
                    \left|
                        \vrtplqt
                    \right\rangle
                    \left\langle
                        \vrtplqt
                    \right|
                    +
                    \left|
                        \hrzplqt{}
                    \right\rangle
                    \left\langle
                        \hrzplqt
                    \right|
                \right)
                .
        \end{equation}
        This model can be defined in terms of a single parameter $t/v$ for which the wave function is
        known exactly at the Rokhsar and Kivelson (RK) point, the point where $t/v =1$. At the RK
        point the wave function is the equal weighted superposition over all dimer configurations
        and known to be in a critical liquid state with powerlaw decaying dimer correlations. This
        model has an exact local $U(1)$ gauge symetry and is known to be described by a $U(1)$ gauge
        thoery.

@@

    In this section we discuss the dimer dimer correlation function
    \begin{equation}
        C_{\mathrm{dimer}} = \langle d_0 d_r \rangle - \langle d_0 \rangle   \langle d_r \rangle   
        ,
    \end{equation}
    where $d_i$ is the ising degrees of freedom on link i.
    In order for the QDPM to exhibit topological order there we must find that it is in a gaped
    dimer liquid phase. It is likely that the QDPM in a liquid phase at the RK point because there 
    we find no evedence of broken symmetries for system sizes up to $32\times32$. As an example of the
    absence of broken symmetries we show
    the dimer correlation function measuered between a single dimer and all other dimers on the
    lattice in Fig.~\ref{fig:fig:dimer_gry_scale_qdpm} for a $16x16$ lattice.
    \subsection{Liquid}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{dimer_gry_scale_qdpm.pdf}
        \caption{Gray scale (actual data) of the dimer corelation between the dimer in the upper
        left corner (colored red) and all other dimers on a $16\times16$ lattice. The 
        dimers in the upperleft corner are colored red because they are off the scale shown at the
        botom of the figure.}
        \label{fig:dimer_gry_scale_qdpm}
    \end{figure}

    The dimer correlations decay exponetialy in the distance between the dimers. 
    In Fig.~\ref{fig:spacial_dmr_cor} we show the dimer correlations between horizontal dimers
    in a vertical column. We find the correlation length to be $HERE$.
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{spacial_dmr_cor.pdf}
        \caption{(Note: This figure was made using equal probabilities of plaquette flips and
        pentamer moves) Spacial correlation function of parallel dimers for the two different
        sublattices distinguished by their color (red, green).}
        \label{fig:spacial_dmr_cor}
    \end{figure}

    \subsection{Comparison to other models???}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating the Gap}
    The exponetial decay of the spacial correlations shown in the previous sections 
    implieds a gap in the spectrum though it is important to determine if the lowest lying
    excitations are vison-like
    or the dimer-like.  The energy gap of the dimer-like and vison-like excitations can be estimated 
    by investigating the dimer and
    vison imaginary time correlations respectivly as outlined by Henley \cite{Henley2004a}. 
    One changlenge in this type of gap estimate is in resloving the correlation functions far enough out
    in imaginary time. In appendix \ref{appendix_gap_err} we discuss our methods used to determine
    that this particular source of systematic error is sufficinetly comparied to the gap
    estimates for different types of excitatoins.
